---
title: "Red_Wine_Quality_Analysis"
author: "Flora Li"
---

What factors will influence the quality of red wine? 
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
library(ggplot2)
library(corrplot)
library(GGally)
library(scales)
library(gridExtra)
library(dplyr)
library(knitr)
library(memisc)
library(tidyr)
library(lattice)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
raw <- read.csv('wineQualityReds.csv')
```

# 1. Introduction
The Red Wine Quality dataset was created by using red and white wine samples. The inputs include objective tests (e.g. PH values) and the output is based on sensory data (median of at least 3 evaluations made by wine experts). Each expert graded the wine quality between 3 (very bad) and 8 (very excellent). This dataset included 1599 observations and effects of 11 different chemical properties.  
```{r, intro}
# dimensions of data
dim(raw)
# variables in data
names(raw)
# print first 10 row
# head(raw,n=10)
```
X is a unique identifier with a integer value. Quality is also an integer value.
All other values are numeric value. In this dataset, we mainly focus on factors 
impacting wine quality, so the quality is dependent variable. 

```{r}
summary(raw)
```
From the summary, we can find out the average quality is 5.6 and the median 
quality is 6. 

# 2. Univariate Plots Section

## Analyzing variable : Quality 
```{r}
table(raw$quality)
```
```{r}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
plot1 <- ggplot(aes(as.factor(quality),fill= factor(quality)), data = raw) + 
  geom_bar() +
  theme_replace() + xlab("quality")+
  scale_fill_brewer(type = 'seq', palette = 14,guide=guide_legend(title = 'Quality'))+
  ggtitle('The Distribution of Quality ')
plot1
```
The plot shows the wine quality basically has a normal distribution. Most of 
the quality are around 5 and 6. 

```{r}
# raw$rating <- ifelse(raw$quality <5, 
#                      'bad', ifelse( raw$quality<7, 'average','good'))
# raw$rating<- ordered(raw$rating, levels = c('bad','average','good'))
# summary(raw$rating)
# qplot(raw$rating)
```

Then we can plot the distribution of all the variables.

```{r}
grid.arrange(qplot(raw$fixed.acidity),
             qplot(raw$volatile.acidity),
             qplot(raw$citric.acid),
             qplot(raw$residual.sugar),
             qplot(raw$chlorides),
             qplot(raw$free.sulfur.dioxide),
             qplot(raw$total.sulfur.dioxide),
             qplot(raw$density),
             qplot(raw$pH),
             qplot(raw$sulphates),
             qplot(raw$alcohol),
             qplot(raw$quality),
             ncol = 4)

             summary(raw$fixed.acidity)
             summary(raw$volatile.acidity)
             summary(raw$citric.acid)
             summary(raw$residual.sugar)
             summary(raw$chlorides)
             summary(raw$free.sulfur.dioxide)
             summary(raw$total.sulfur.dioxide)
             summary(raw$density)
             summary(raw$pH)
             summary(raw$sulphates)
             summary(raw$alcohol)
             summary(raw$quality)


```

From the plots above, we can find out: 
1.Density and ph are basically normal distributed.
2.Fixed and volatile acidity, sulfur dioxides, sulphates, and alcohol are right 
skewed distributions, with means their means are larger than median.
3.Residual sugar and chlorides are also right skewed and they have extreme
outliers.

### Rescale features

Skewed and long tail data can be transformed toward more normally distribution 
by taking square root or log function. Since residual sugar have lower 
correlation with quality (see Bivariate Plots Section), I just log-transformed 
the sulphates to compare the differences.

```{r,log sulphates}
# sulphates 

p1<-ggplot(aes(x = sulphates), data = raw)+
        geom_histogram(colour="black", fill="pink" 
                      ,binwidth = 0.1)

# log sulphates
raw$log_sulphates <- log(raw$sulphates)

p2<-ggplot(aes(x = log_sulphates), data = raw)+
          xlab("log of sulphates")+
          geom_histogram(colour="black", fill="pink", 
                         binwidth = 0.1)

grid.arrange(p1,p2,ncol=1)

```
For the first plot, we can see feature Sulphates have really large numbers and 
is right skewed. After log transformation, the sulphates is more normal distributed. 


# Univariate Analysis

### What is the structure of your dataset?
```{r}
# data structure 
str(raw)
```

### What is/are the main feature(s) of interest in your dataset?
The quality of red wine is the main feature of interest in this dataset.

### What other features in the dataset do you think will help support your \
### investigation into your feature(s) of interest?
From the plot above, I think the variables related to acidity (fixed, volatile,
citric.acid and pH) will influence the taste of red wine. Citric acid provides 
freshness taste to wines. Thus, better red wine would contain higher citric acid
rate. Volatile Acidity provides unpleasant and vinegar taste to wines so the 
lower the volatile acidity, the better the red wine is. Besides, the percentage 
of alcohol will also influence the quality of wine. With higher percent of
alcohol, the red wine would have better quality. Residual.sugar dictates how 
sweet a red wine is and will also have an influence in taste.

### Did you create any new variables from existing variables in the dataset?
No. I didn'd create any new variables.

### Of the features you investigated, were there any unusual distributions? \
### Did you perform any operations on the data to tidy, adjust, or change the form \
### of the data? If so, why did you do this?

For the variable "Quality", there are no unusual distributions. It's basically 
normal distributed. For the other features such as sulphates and acidity,they 
are long-tailed.I tried log transformation because the it will be more normally 
distributed after log transformation.
 


# Bivariate Plots Section

### Correlation Matrix
```{r echo=FALSE, Bivariate_Plots}
correlation <- cor(raw[2:13])
round(correlation,digits = 3)
corrplot(correlation)
```

From the result, we can find out that quality has higher positive correlation\
with alcohol(0.47),sulphates(0.25),citric acid(0.22) and fixed acid(0.124). It \
also has higher negative correlation with valatile acidity(-0.39),total sulfur
dioxide(-0.185) and density(-0.175).

### Boxplot of Quality 

```{r}
quality_plot <- function (x, y, ylab) {
return (ggplot(data = raw, aes_string(x,y)) +
geom_boxplot(fill = 'pink') +
xlab ('quality') + ylab(ylab))

}

grid.arrange( quality_plot( 'factor(quality)', 'fixed.acidity',
                            'fixed.acidity(g/dm^3)'), 
quality_plot('factor(quality)', 'volatile.acidity', 'volatile.acidity(g/dm^3)'),
quality_plot('factor(quality)', 'citric.acid', 'citric.acid (g / dm^3)'),
quality_plot('factor(quality)', 'residual.sugar', 'residual.sugar (g / dm^3)'),
quality_plot('factor(quality)', 'chlorides', 'chlorides (g / dm^3)'),
quality_plot('factor(quality)', 'free.sulfur.dioxide', 
             'free.sulphur.dioxide (g / dm^3)'),
quality_plot('factor(quality)', 'total.sulfur.dioxide',
             'total.sulphur.dioxide (g / dm^3)'),
quality_plot('factor(quality)', 'density', 'density (g/cm^3)'),
quality_plot('factor(quality)', 'pH', 'pH'),
quality_plot('factor(quality)', 'sulphates', 'sulphates (g/dm^3)'),
quality_plot('factor(quality)', 'alcohol', 'alcohol (volume %)'),
 
ncol= 4,
top = "Quality & All Features (Boxplot)")

```

Based on the plots above, we can infer good wines have the following attribures:
1.Lower fixed acidity, volatile acidity, density and ph.All these features would 
make the wine taset bad so
2.Higher alcohol,sulphates and citric acidity.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
### investigation. How did the feature(s) of interest vary with other features in \
### the dataset?

The top 5 features correlated with red wine quality:
1. Density (-0.496)
2. Alcohol (0.476)
3. Volatile acidity (-0.39)
4. Sulphates (0.251)
5. Citric acid (0.22)
From this result we can find out, good red wine tend to have lower density and 
valatile acidity and higher alcohol, sulphates and citric acid.

```{r}
cor.test(raw$quality, raw$density,method='pearson') 
cor.test(raw$quality, raw$alcohol,method='pearson')
cor.test(raw$quality, raw$volatile.acidity,method='pearson') 
cor.test(raw$quality, raw$sulphates,method='pearson') 
cor.test(raw$quality, raw$citric.acid,method='pearson') 
```
From the Pearson's correlation test, we can find out density is less correlated with quality
compared to other features.

It's also observed free sulfur dioxide and residual sugar don't have much effect 
on the quality of red wine.


### Did you observe any interesting relationships between the other features \
### (not the main feature(s) of interest)?

Observations about relationship between supporting variables:

1.All the acid features(fixed acidity, volatile acidity and citric acidity) are
highly correlated.Because volatile Acidity provides unpleasant and vinegar taste
to wines, the lower the volatile acidity, the better the wine is.While citric 
acid provides freshness taste to wines. Thus, better red wine would contain
higher citric acid rate. In this dataset, I mainly focus on volatile acidity and 
citric acid because compared to fixed acidity, they seems have higher correlation
with wine quality.
2.Acid features have higher correlation with ph.This makes sense because 
any ph number less than 7 is considered an acid.
3.Acid feature have higher correlation with sulphates and density.
4.Density has higher positive correlation with residual sugar and chlorides.

### What was the strongest relationship you found?
The relationship between total sulfur dioxide and free sulfur dioxide.The 
correlation is 0.668.


----
# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}
ggplot(raw, aes(y=alcohol,x=quality,color = volatile.acidity ) )+
        geom_jitter(alpha=1/2,size=3)+
        xlab("wine quality") + ylab("alcohol") +
        scale_colour_gradient2(low="blue", high="red",
                               mid="pink",midpoint = 0.8)

```

We can add another feature, the log scale of sulphates.In order to visualize 
the result, we can classify the quality into 3 categories, good, average and bad,
and name this new feature as rating.

```{r}
# add new feature rating
raw$rating <- ifelse(raw$quality <5, 
                    'bad', ifelse( raw$quality<7, 'average','good'))
raw$rating<- ordered(raw$rating, levels = c('bad','average','good'))

plot3<- ggplot(raw, aes(x=alcohol,y=volatile.acidity)) +
      geom_point(aes(color =  log_sulphates),alpha=1/2,size=2)+
      xlab("alcohol") + ylab("volatile acidity") +
      scale_colour_gradientn(colours = terrain.colors(10)) +
facet_grid(rating ~ .) +
   ggtitle("Wine Rating vs Log_sulphates vs violatile acidity vs alcohol")
plot3

```


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
### investigation. Were there features that strengthened each other in terms of \
### looking at your feature(s) of interest?
For the first plot, we can observe higher quality wine have higher alcohol and 
lower volatile acidity. For the second plot, we can see higher quality wine have
higher alcohol (x-axis),lower volatile acidity (y-axis) and higher sulphates (hue).

### Were there any interesting or surprising interactions between features?
It's interesting to find out that in second plot if volatile acidity is lower,
sulphates will be higher.

------

# Final Plots and Summary

### Plot One: The histogram of "Quality"
```{r echo=FALSE, Plot_One}
plot1
```

### Description One
The wine quality basically has a normal distribution. Most of 
the quality are 5 and 6. Followed by quality 7 and quality 4. Quality 8 followed
by quality 3 are the least available. 




### Plot Two
```{r echo=FALSE, Plot_Two}
quality_plot <- function (x, y, ylab) {
return (ggplot(data = raw, aes_string(x,y)) +
geom_boxplot(fill = 'pink') +
xlab ('quality') + ylab(ylab))

}

grid.arrange( quality_plot( 'factor(quality)', 'fixed.acidity',
                            'fixed.acidity(g/dm^3)'), 
quality_plot('factor(quality)', 'volatile.acidity', 'volatile.acidity(g/dm^3)'),
quality_plot('factor(quality)', 'citric.acid', 'citric.acid (g / dm^3)'),
quality_plot('factor(quality)', 'residual.sugar', 'residual.sugar (g / dm^3)'),
quality_plot('factor(quality)', 'chlorides', 'chlorides (g / dm^3)'),
quality_plot('factor(quality)', 'free.sulfur.dioxide', 
             'free.sulphur.dioxide (g / dm^3)'),
quality_plot('factor(quality)', 'total.sulfur.dioxide',
             'total.sulphur.dioxide (g / dm^3)'),
quality_plot('factor(quality)', 'density', 'density (g/cm^3)'),
quality_plot('factor(quality)', 'pH', 'pH'),
quality_plot('factor(quality)', 'sulphates', 'sulphates (g/dm^3)'),
quality_plot('factor(quality)', 'alcohol', 'alcohol (volume %)'),
 
ncol= 4,
top = "Quality & All Features (Boxplot)")
```

### Description Two
Based on the plots above, we can infer good wines have the following attribures:
1.Lower fixed acidity, volatile acidity, density and ph.All these features would 
make the wine taset bad so
2.Higher alcohol,sulphates and citric acidity.

### Plot Three
```{r echo=FALSE, Plot_Three}
plot3
```

### Description Three
Higher quality wine have higher alcohol (x-axis),lower volatile acidity (y-axis)
and higher sulphates (hue).


------

# Reflection

From this project, I find out that higher quality red wine have higher alcohol 
,lower volatile acidity and higher sulphates.I was surprising to found out that
some chemical parameters such as residual sugar, chlorides, free sulfur dioxide 
and total sulfur dioxide did not have much impact on the quality of red wine.

For the third plot,at first, I didn't use log transformation to deal with the 
sulphates data. The result looked so confusing because it's basically all green 
dots in the plot.After I use the log transformation, the trend looks more obvious.

The analysis could have been performed better if there was more data on other
qualities of red wines available. The results would be more accurate if more data
was availabe. It's also possible the results would be totally different with more
data.

For future explorations I can combine the red wine dataset with the dataset for
white wine to compare analyze the differences. 


References:
Correlation Matrix
https://rstudio-pubs-static.s3.amazonaws.com/240657_5157ff98e8204c358b2118fa69162e18.html

http://genomicsclass.github.io/book/pages/dplyr_tutorial.html

https://github.com/Dalaska/Udacity-Red-Wine-Quality/blob/master/redwine_final.rmd

https://github.com/SThornewillvE/Udacity-Project---Exploring-Wine-Data/blob/master/investigating-a-dataset.Rmd

https://github.com/anilsai/Explorartory-data-Analytics_-R-_Udacity-_Wine-Data/blob/master/Wine_quality_Anil%20Bodepudi_new.Rmd

https://github.com/jeswingeorge/EDA-project---wineQualityReds/blob/master/projecttemplate.rmd

